---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

***Introduction***

There are persistent problems and injustices in the United States around educational and income inequality. For our research project, we are going to examine what factors influence career outcomes (one's education and income) by looking at data from over 3,000 counties in the United States. In many cases, these education and income are very interrelated for families and counties causing or exacerbating a vicious cycle that can be hard to break. In America, there are many root causes that contribute to such a cycle, and it is important they are clearly researched and evaluated, so changes can start to be made to populations and counties that need them most. There are many factors that can influence one's career outcomes, but some particularly important ones are education level, ethnicity, gender, and state. According to the APLU, people who hold college degrees are twice as likely to get a job and make around an extra \$1.2 million over the course of their life (APLU). Black or Latino families make around 50% of what white households do, according to the Federal Reserve Board (Federal Reserve Board). The gender pay gap still persists, with women earning an average of 82% of what men do (PEW Research Center). On the state level, various factors such as funding per county and class disparities can impact income inequality as well. There are intersections in everyone's identities and therefore the counties identities as well, and when combined, can have even greater impacts on career outcomes, so knowing which specific identities are most vulnerable is also a key step to making actionable change. We believe counties with residents with higher levels of education, a higher white population, more males, and live in states with greater funding and job opportunities will have better career outcomes. Since this data has been gathered by the US Census Bureau, a government organization, there should be no ethical concerns. They adhere to strict privacy policy guidelines to ensure a confidential, respectful, and transparent experience for participants when collecting data (United States Census Bureau).

<https://www.census.gov/about/policies/privacy/data_stewardship/our_privacy_principles.html>

<https://www.aplu.org/our-work/4-policy-and-advocacy/publicuvalues/employment-earnings/#:~:text=The%20evidence%20that%20a%20college,on%20average%20over%20their%20lifetime>.

<https://www.federalreserve.gov/econres/notes/feds-notes/wealth-inequality-and-the-racial-wealth-gap-20211022.html#:~:text=In%20the%20United%20States%2C%20the,percent%20as%20much%20net%20wealth>.

<https://www.pewresearch.org/fact-tank/2023/03/01/gender-pay-gap-facts/#:~:text=The%20gender%20gap%20in%20pay,%2D%20and%20part%2Dtime%20workers>.

\
***Data***

The dataset we are analyzing is from The Collection of Really Great, Interesting, Situated Datasets (CORGIS) Project. Furthermore, it was collected using 2010 to 2019 data from the United States Census Bureau. It should be noted that some of the estimates came from sample data. The dataset includes a wide variety of observations for various counties in the United States. In order to best serve our research question, we will focus on the statistics such as education level, ethnicity, age, gender, and state. The data is largely qualitative and in terms of what percentage of the county fits a certain description such as being a high school graduate or white.\

***Methodology***

```{r}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
library(usmap)
library(geofacet)
```

```{r}
#| label: load-data-2
#| message: false
#| warning: false

counties <- read_csv("data/county_demographics.csv")

glimpse(counties)
```

```{r}
#| label: linreg_ethnicity
#| eval: false
county_fit <- linear_reg() |>
  set_engine("lm") |>
  fit(`Income.Median Houseold Income` ~ 
        `Ethnicities.White Alone	 not Hispanic or Latino` + 
        `Ethnicities.Hispanic or Latino`, 
      data = counties, family = "binomial") 
```

```{r}
#| label: states
#| message: false
#| warning: false

statepop <- statepop

state_data <- counties |>
  group_by(State) |>
  summarize(median_income = median(`Income.Median Houseold Income`),
            median_education = median(`Education.Bachelor's Degree or Higher`))

maps_dataframe <- full_join(statepop, state_data, by = c("abbr" = "State")) 

plot_usmap(data = maps_dataframe,
           values = "median_income",
           labels = TRUE) +
  labs(fill = "Median Household Income", 
       title = "Household Incomes Per State") +
  scale_fill_viridis_c() +
  theme(legend.position = "bottom") 

plot_usmap(data = maps_dataframe,
           values = "median_education",
           labels = TRUE) +
  labs(fill = "Median Education Levels", 
       title = "Education Levels Per State") +
  scale_fill_viridis_c() +
  theme(legend.position = "bottom") 
```

***Results***

According to the maps, education and income levels appear to be the lowest in the southern regions, especially in the southeast. They tend to be the highest in the northeast and west.
